# Mandatory Inputs

```{r}
library(tidyverse)
library(ggplot2)
library(patchwork)
library(knitr)
library(cowplot)
```

```{r}
# csv that contains the data of the study
df = read.csv("/run/user/1000/gvfs/smb-share:server=upcourtinenas,share=cervical/CERVICAL_ID/connectome_analysis/mutiple_brain_analysis/all_brains.csv")
# csv that containes the metadato of the samples
df_meta = read.csv("/run/user/1000/gvfs/smb-share:server=upcourtinenas,share=cervical/CERVICAL_ID/connectome_analysis/mutiple_brain_analysis/all_brains_meta.csv")
# Dir where to save all the plots
base_dir = "/run/user/1000/gvfs/smb-share:server=upcourtinenas,share=cervical/CERVICAL_ID/Connectome_analysis/Mutiple_brain_analysis" 
# Csv with ROI nomenclature and full name
df_roi_names = read_csv("/home/gabri/Desktop/intership_neurorestore_pipeline/assets/leaves_regions_atlas.csv")
```

# Evolution csv

```{r}
# GOAL:
  # Take a region of injection (ex "DR)
  # Create a csv where
    # Each row is a ROI
    # Each column is the mean(or std) densoty for the ROI at a specific timepoint
#to otbian smethign like this:
#ROI           Mean_Uninjured   Std_Uninjured   Mean_8 weeks   Std_8 weeks
#Left: FRP1    0.000040         0.000057        0.000040       0.000057
#Right: FRP1   0.000094         0.000133        0.000094       0.000133
#Left: FRP2/3  0                0               0              0


# Regions of injection to plot (even more than one)
regions_injection = c("DR")

for (region in regions_injection) {  # Iterate over the region_injection
  
  print(region)
  
  #Crete df to fill. Intialize with only the roi names
  df_evolution <- df %>%
    select(ROI) %>%
    distinct()#take only unqiue, indeed orignal df, have the roi repteas as many times as the brains
  
  # Create temp copy of df
  count_data = df
  meta_data = df_meta
  
  # Select mice injected in specific region
  count_data = count_data[count_data$`Region.Injection` == region,]
  
  # Change the shape of the df
  count_data <- count_data %>%
    select(ROI, `Cell.Density`, `Brain.ID`) %>%
    pivot_wider(names_from = `Brain.ID`, values_from = `Cell.Density`)
  
  # Select mice injected in specific region
  meta_data = meta_data[meta_data$`Brain.ID` %in% colnames(count_data),]
  
  # Take the possbile values fo timepoints
  timepoints = rev(unique(df_meta$TimePoint)) #rev for invert order
  
  #make a plot for each time point
  for (timepoint in timepoints){
    
    temp_meta = meta_data[meta_data$TimePoint == timepoint, ] #Take only the mouse of these time point
    temp_data = count_data[, colnames(count_data) %in% c(temp_meta$`Brain.ID`,"ROI")] #Take only the relative brains
    
    #Calculate mean and std
    temp_data_no_first_col <- temp_data[, -1] #remove first column that are the nam eof the ROI
    row_means <- apply(temp_data_no_first_col, 1, mean)
    row_means[is.nan(row_means)] = 0
    row_stds <- apply(temp_data_no_first_col, 1, sd)
    row_stds[is.nan(row_stds)] = 0
                     
    # Add columns to df
    temp_data <- temp_data %>%
      mutate(row_mean = row_means, row_std = row_stds)
    
    #print(temp_data_sorted)
    
    #Add columns to evolution_df
    new_columns <- temp_data %>%
      select(row_mean, row_std) %>%
      rename_with(~paste0(c("Mean ", "Std "), timepoint), everything())
    df_evolution <- df_evolution %>%
      bind_cols(new_columns)

  }

  
  #Save the csv
  dir_images = file.path(base_dir, "images", region)
  #print(dir_images)
  if (!dir.exists(dir_images)) {
    dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
  }
  
  print(df_evolution)
  
  write_csv(df_evolution, paste0(dir_images, "/evolution_timeline.csv"))

}

```

# Plots

```{r}
n_roi_displayed = 10
all_region_injection = unique(df$Region.Injection)
full_name_plot = TRUE # True if you want full name on the x-axis of the plots


#+++++++++++++++++++++++++
# Function to calculate the mean and the standard deviation
  # for each group
#+++++++++++++++++++++++++
# data : a data frame
# varname : the name of a column containing the variable
  #to be summariezed
# groupnames : vector of column names to be used as
  # grouping variables
data_summary <- function(data, varname, groupnames){
  require(plyr)
  summary_func <- function(x, col){
    c(mean = mean(x[[col]], na.rm=TRUE),
      std = sd(x[[col]], na.rm=TRUE))
  }
  data_sum<-ddply(data, groupnames, .fun=summary_func,
                  varname)
  data_sum <- rename(data_sum, c("mean" = varname))
 return(data_sum)
}
```

## Barplot Density Different X-Axis

**GOAL:**

1.  **Define Regions of Injection:**
    -   Specify the regions to be analyzed (e.g., "DR").
2.  **Read Evolution Data:**
    -   For each specified region, load the data from a CSV file containing evolution metrics.
3.  **Find Timepoints:**
    -   Extract and reverse the order of unique timepoints for the region.
4.  **Create Timepoint Plots:**
    -   For each timepoint:
        -   Select the columns related to the mean and standard deviation for the timepoint.
        -   Identify and sort the top `n_roi_displayed` ROIs based on mean density.
        -   Generate a bar plot showing the mean density with error bars.
        -   Optionally, add a text table with ROI names.
5.  **Combine and Save Plots:**
    -   Combine the plots for all timepoints into a single figure.
    -   Save the combined plot as a PDF file in a specified directory for each region.

```{r}
# GOAL:
  # Take a region of injection (ex "DR)
  # For each timePoint (ex. "Uninjured")
    # For each ROI 
      # Calculate the mean density across the mouse of the timepoit choosen
  # Take the most "n_roi_displayed" dense region for each time point
    # (so each timepint will have a differt set of ROI)
  # Create a barplot for each time point
    #Put it in the same figure

# Regions of injection to plot (even more than one)
regions_injection = c("DR")

for (region in regions_injection) {  # Iterate over the region_injection
  
  print(region)
  
  # Read df evolution
  df_evolution_region = read_csv(file.path(base_dir, "images", region, "evolution_timeline.csv"))
  
  # Find timepoints of this injection region
  timepoints = df_meta %>% filter(Region.Injection == region) %>% select(TimePoint)
  timepoints = rev(unique(timepoints$TimePoint))
  
  # TimePoint plot
  timepoints_plots = list()
  
  #make a plot for each time point
  for (timepoint in timepoints){
    
    # Name of the mean column and Std in the df
    means_col_name <- paste0("Mean ", timepoint)
    stds_col_name <- paste0("Std ", timepoint)
    
    # Suset to only columns relative to this timeponr
    t = df_evolution_region %>% select(ROI, means_col_name, stds_col_name)
    
    #change name col
    names(t)[names(t) == means_col_name] = "means"
    names(t)[names(t) == stds_col_name] = "stds"

    # Sort the data frame by 'mean' in descending order
    # Select the first most dense regions
    t_sorted <- t %>%
      arrange(desc(means)) %>%
      slice_head(n = n_roi_displayed)
    
    # to force the ROI to be displayed on the x axis on the same order of the most dense
    t_sorted <- t_sorted %>%
      mutate(ROI = factor(ROI, levels = ROI))
    
    #print(t_sorted)
    
    # Create the bar plot with error bars
    plot = ggplot(t_sorted, aes(x = ROI, y = means)) +
      geom_bar(stat = "identity", fill = "lightblue", color = "blue") +
      geom_errorbar(aes(ymin = means - stds, ymax = means + stds), 
                    width = 0.2, color = "red") +
      labs(title = timepoint, x = paste0("First most dense ", n_roi_displayed, " ROI"), y = "Density") +
      coord_cartesian(ylim = c(0, 0.02)) +  # Set y-axis limit
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Tilt x-axis labels
    
        # Add Table with ROI names if Needed
    if (full_name_plot){
      # Filter and merge
      roi_acronyms <- t_sorted$ROI
      roi_names <- df_roi_names %>%
        filter(Acronym %in% roi_acronyms) %>%
        rename(ROI = Acronym)
      
      t_sorted <- merge(t_sorted, roi_names, by = "ROI", all.x = TRUE)
      
      # Create concatenated column
      t_sorted <- t_sorted %>%
        mutate(Acronym_Name = paste(ROI, Name, sep = " --> "))
      
      # Create the text_lines with each entry left-aligned
      text_lines <- paste(t_sorted$Acronym_Name, collapse = "\n")
      
      # Create a plot with text annotation
      text_plot <- ggdraw() + 
        draw_label(text_lines, size=8, 
                   hjust=0, vjust=0.5, # how to justify all the text, ex. all aligned to the left
                   x=0.2, y=0.6 #coodinate of the text
                                # x left/right, y up/down
                   )+
        theme(plot.margin = margin(0, 0, 0, 0))
      
      # Combine the plot and text side by side
      plot = plot_grid(plot, text_plot, ncol=1, rel_heights = c(3, 1))

    }

    #print(plot)
    
    #put plot in list
    timepoints_plots[[timepoint]] <- plot
  }

  # Print the combined plot
  combined_plot <- wrap_plots(timepoints_plots, ncol = length(timepoints_plots))
  
  print(combined_plot)
  
  #Save the figure
  dir_images = file.path(base_dir, "images", region)
  #print(dir_images)
  if (!dir.exists(dir_images)) {
    dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
  }
  ggsave(file.path(dir_images, paste0("barplots_density_different_ROI_", n_roi_displayed, ".pdf")), combined_plot, width = 10, height = 6)

}

```

## Barplot Density Same X-Axis (Unijured as baseline)

**GOAL:**

1.  **Define Regions of Injection:**
    -   Specify the regions to be analyzed (e.g., "DR").
2.  **Read Evolution Data:**
    -   For each specified region, load the data from a CSV file containing evolution metrics.
3.  **Find Timepoints:**
    -   Extract and reverse the order of unique timepoints for the region.
4.  **Create Timepoint Plots:**
    -   For each timepoint:
        -   **Uninjured Timepoint:** Identify and save the most dense ROIs based on mean density.
        -   **Other Timepoints:** Use the previously identified most dense ROIs and sort by the same order.
        -   Generate bar plots showing mean density with error bars.
        -   Optionally, add a text table with ROI names.
5.  **Combine and Save Plots:**
    -   Combine the plots for all timepoints into a single figure.
    -   Save the combined plot as a PDF file in a specified directory for each region.

```{r}
# GOAL:
  # Take a region of injection (ex "DR)
  # For each timePoint (ex. "Uninjured")
    # For each ROI 
      # Calculate the mean density across the mouse of the timepoit choosen
  # Take the most "n_roi_displayed" dense region for Uninjured case
    # show the same ROI for all the differt timepoints
  # Create a barplot for each time point
    #Put it in the same figure

# Regions of injection to plot (even more than one)
regions_injection = c("DR")

for (region in regions_injection) {  # Iterate over the region_injection
  
  print(region)
  
  # Read df evolution
  df_evolution_region = read_csv(file.path(base_dir, "images", region, "evolution_timeline.csv"))
  
  # Find timepoints of this injection region
  timepoints = df_meta %>% filter(Region.Injection == region) %>% select(TimePoint)
  timepoints = rev(unique(timepoints$TimePoint))
  
  # TimePoint plot
  timepoints_plots = list()
  
  #Name most dense region of uninjured (baseline)
  roi_dense_uninjured = list()
  
  #make a plot for each time point
  for (timepoint in timepoints){
    
    if(timepoint == "Uninjured"){ #Let's search the most dense regions
      
      # Name of the mean column and Std in the df
      means_col_name <- paste0("Mean ", timepoint)
      stds_col_name <- paste0("Std ", timepoint)
      
      # Suset to only columns relative to this timeponr
      t = df_evolution_region %>% select(ROI, means_col_name, stds_col_name)
      
      #change name col
      names(t)[names(t) == means_col_name] = "means"
      names(t)[names(t) == stds_col_name] = "stds"
  
      # Sort the data frame by 'mean' in descending order
      # Select the first most dense regions
      t_sorted <- t %>%
        arrange(desc(means)) %>%
        slice_head(n = n_roi_displayed)
      
      # to force the ROI to be displayed on the x axis on the same order of the most dense
      t_sorted <- t_sorted %>%
        mutate(ROI = factor(ROI, levels = ROI))
  
      #print(t_sorted)
      
      #Save most dense regions
      roi_dense_uninjured = as.character(t_sorted$ROI)
    }
    else{#Use the most dense ROI of uninjured
      
      # Name of the mean column and Std in the df
      means_col_name <- paste0("Mean ", timepoint)
      stds_col_name <- paste0("Std ", timepoint)
      
      # Suset to only columns relative to this timeponr
      t = df_evolution_region %>% select(ROI, means_col_name, stds_col_name)
      
      #change name col
      names(t)[names(t) == means_col_name] = "means"
      names(t)[names(t) == stds_col_name] = "stds"
      
      # Take only ROI of unibjured
      # Sort BY THE SAME ORDER OF UNIJURED (NOT DESCENDING)
      # Select the SAME ROI OF UNIJURED !!!!!!
      t_sorted <- t %>%
        filter(ROI %in% roi_dense_uninjured) %>%
        mutate(ROI = factor(ROI, levels = roi_dense_uninjured)) #TO CHNAGE THE ORDER
      
      #print(t_sorted)
      
    }
    
    # Create the bar plot with error bars
    plot = ggplot(t_sorted, aes(x = ROI, y = means)) +
      geom_bar(stat = "identity", fill = "lightblue", color = "blue") +
      geom_errorbar(aes(ymin = means - stds, ymax = means + stds), 
                    width = 0.2, color = "red") +
      labs(title = timepoint, x = paste0("First most dense ", n_roi_displayed, " ROI"), y = "Density") +
      coord_cartesian(ylim = c(0, 0.02)) +  # Set y-axis limit
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Tilt x-axis labels
    
    # Add Table with ROI names if Needed
    if (full_name_plot){
      # Filter and merge
      roi_acronyms <- t_sorted$ROI
      roi_names <- df_roi_names %>%
        filter(Acronym %in% roi_acronyms) %>%
        rename(ROI = Acronym)
      
      t_sorted <- merge(t_sorted, roi_names, by = "ROI", all.x = TRUE)
      
      # Create concatenated column
      t_sorted <- t_sorted %>%
        mutate(Acronym_Name = paste(ROI, Name, sep = " --> ")) 
      
      # Create the text_lines with each entry left-aligned
      text_lines <- paste(t_sorted$Acronym_Name, collapse = "\n")
      
      # Create a plot with text annotation
      text_plot <- ggdraw() + 
        draw_label(text_lines, size=8, 
                   hjust=0, vjust=0.5, # how to justify all the text, ex. all aligned to the left
                   x=0.2, y=0.6 #coodinate of the text
                                # x left/right, y up/down
                   )+
        theme(plot.margin = margin(0, 0, 0, 0))
      
      # Combine the plot and text side by side
      plot = plot_grid(plot, text_plot, ncol=1, rel_heights = c(3, 1))

    }

    #print(plot)
    
    #put plot in list
    timepoints_plots[[timepoint]] <- plot
    
  }

  # Print the combined plot
  combined_plot <- wrap_plots(timepoints_plots, ncol = length(timepoints_plots))
  print(combined_plot)
  
  #Save the figure
  dir_images = file.path(base_dir, "images", region)
  #print(dir_images)
  if (!dir.exists(dir_images)) {
    dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
  }
  ggsave(file.path(dir_images, paste0("barplots_density_same_ROI_", n_roi_displayed, ".pdf")), combined_plot, width = 10, height = 6)


  
}

```

## Barplot Density Multiple columns (Unijured as baseline)

```{r}
# Regions of injection to plot (even more than one)
regions_injection = c("DR")

for (region in regions_injection) {  # Iterate over the region_injection
  
  print(region)
  
  d = data_summary(df, varname="Cell.Density", 
                    groupnames=c("TimePoint", "Region.Injection", "ROI")) %>%
      dplyr::rename(mean = `Cell.Density`) %>%
      filter(Region.Injection == region)
      
  
  # Take the most dense region of Uninjured
  roi_dense_uninjured  = d %>%
      filter(TimePoint == "Uninjured") %>%
      arrange(desc(mean)) %>%  # Sort in descending order
      select(ROI) %>%  # Select only the ROI column
      slice_head(n = n_roi_displayed) %>%
      mutate(ROI = factor(ROI, levels = ROI))
  roi_dense_uninjured = as.character(roi_dense_uninjured$ROI)
  
  df_temp = d %>% filter(ROI %in% roi_dense_uninjured)
  
  plot = ggplot(data = df_temp, 
       aes(x = ROI, y = mean, fill = TimePoint)) + 
       geom_bar(stat = 'identity', position = 'dodge') +
       geom_errorbar(aes(ymin = mean - std, ymax = mean + std), position = position_dodge(width = 0.7), width = 0.2) +  # Error bars
       labs(title = timepoint, x = paste0("First most dense ", n_roi_displayed, " ROI"), y = "Density") +
       coord_cartesian(ylim = c(0, 0.02)) +  # Set y-axis limit
       theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Tilt x-axis labels
  
  print(plot)
  
  
      
      
}
```

## Log(FC) barplots

**GOAL:**

1.  **Define Regions of Injection:**
    -   Specify the regions to be analyzed (e.g., "DR").
2.  **Read Evolution Data:**
    -   For each specified region, load the data from a CSV file containing evolution metrics.
3.  **Generate Timepoint Comparisons:**
    -   Create combinations of two timepoints for comparison (e.g., "Uninjured" vs. "Injured").
4.  **Calculate and Plot Log Fold Change (logFC):**
    -   For each timepoint combination, calculate the logFC between the two timepoints.
    -   Handle zero values and missing data appropriately.
    -   Identify and sort the top and bottom ROIs based on logFC.
5.  **Create and Customize Plots:**
    -   Generate bar plots showing the logFC for top and bottom ROIs.
    -   Optionally add a text table with ROI names.
    -   Arrange the plots side by side for all timepoint comparisons.
6.  **Save Plots:**
    -   Save the combined plots as a PDF file in a specified directory for each region.

```{r}


# Regions of injection to plot (even more than one)
regions_injection = c("DR")

# Read df evolution

for (region in regions_injection) {  # Iterate over the region_injection
  
  print(region)
  
  # Read df evolution
  df_evolution_region = read_csv(file.path(base_dir, "images", region, "evolution_timeline.csv"))
  
  # Create all possible combination of 2 timepoits
  timepoints = df_meta %>% filter(Region.Injection == region) %>% select(TimePoint)
  timepoints = rev(unique(timepoints$TimePoint))
  combinations_timepoints = combn(timepoints, 2, simplify = FALSE)
  
  # Comb plot
  comb_plots = list()
  
  for(comb in combinations_timepoints){
    #print(comb)
    
    # Names of the 2 classes in this co,bination
    baseline_class = comb[1] #Nb index start from 1
    treatment_class = comb[2]
    
    # Name of the mean column in the df
    baseline_col <- paste0("Mean ", baseline_class)
    treatment_col <- paste0("Mean ", treatment_class)
    
    baseline_data <- as.numeric(df_evolution_region[[baseline_col]])
    treatment_data <- as.numeric(df_evolution_region[[treatment_col]])
    
    # Handle division by zero and NA
    # Replace zeros in baseline data with a small value to avoid division by zero
    baseline_data[baseline_data == 0] <- .Machine$double.eps
    
    # Calculate log FC, handling division by zero and NA
    log_fc <- log2(treatment_data / baseline_data)
    
    # Replace NA values with 0
    log_fc[is.na(log_fc)] <- 0
    # Replace Inf with 0
    log_fc[is.infinite(log_fc)] <- 0

    # Combine with ROI in a new data frame
    df_temp <- df_evolution_region %>%
      select(ROI) %>%
      mutate(logFC = log_fc)
    
    # Sort by row_mean in descending order and select the top n rows
    top_rows <- df_temp %>%
      arrange(desc(logFC)) %>%
      slice_head(n = n_roi_displayed)
    
    # Sort by row_mean in ascending order and select the bottom n rows
    bottom_rows <- df_temp %>%
      arrange(logFC) %>% #ascending
      slice_head(n = n_roi_displayed) %>%
      arrange(desc(logFC)) #the ramiang row i want the form biggest tos amllest

    # Combine the top and bottom rows
    combined_rows <- bind_rows(top_rows, bottom_rows)
    
    # to force the ROI to be displayed on the x axis on the same order of the most dense
    combined_rows <- combined_rows %>%
      mutate(ROI = factor(ROI, levels = ROI))
        #you are defining the levels of the factor in the same order as the values currently appear in the ROI column of combined_rows
  
    # Create the bar plot with error bars
    plot = ggplot(combined_rows, aes(x = ROI, y = logFC)) +
      geom_bar(stat = "identity", fill = "lightblue", color = "blue") +
      labs(title = paste0("Baseline: ", comb[1], "\nTreatment: ", comb[2]), x = paste0("First biggest ", n_roi_displayed, " log(FC)"), y = "lof(FC)") +
      #coord_cartesian(ylim = c(0, 0.02)) +  # Set y-axis limit
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Tilt x-axis labels

    # Add Table with ROI names if Needed
    if (full_name_plot){
      
      t_sorted = combined_rows
      
      # Filter and merge
      roi_acronyms <- t_sorted$ROI
      roi_names <- df_roi_names %>%
        filter(Acronym %in% roi_acronyms) %>%
        rename(ROI = Acronym)
      
      t_sorted <- merge(t_sorted, roi_names, by = "ROI", all.x = TRUE)
      
      # Create concatenated column
      t_sorted <- t_sorted %>%
        mutate(Acronym_Name = paste(ROI, Name, sep = " --> ")) 
      
      # Create the text_lines with each entry left-aligned
      text_lines <- paste(t_sorted$Acronym_Name, collapse = "\n")
      
      # Create a plot with text annotation
      text_plot <- ggdraw() + 
        draw_label(text_lines, size=8, 
                   hjust=0, vjust=0.5, # how to justify all the text, ex. all aligned to the left
                   x=0.2, y=0.5 #coodinate of the text
                                # x left/right, y up/down
                   )+
        theme(plot.margin = margin(0, 0, 0, 0))
      
      # Combine the plot and text side by side
      plot = plot_grid(plot, text_plot, ncol=1, rel_heights = c(2, 1))

    }
    
    #print(plot)
    
    #put plot in list
    comb_plots[[paste0(comb[1], comb[2])]] <- plot
  }

  # Print the combined plot
  combined_plot <- wrap_plots(comb_plots, ncol = length(combinations_timepoints))
  print(combined_plot)
  
  #Save the figure
  dir_images = file.path(base_dir, "images", region)
  #print(dir_images)
  if (!dir.exists(dir_images)) {
    dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
  }
  ggsave(file.path(dir_images, paste0("barplots_logFC_", n_roi_displayed, ".pdf")), combined_plot, width = 10, height = 8)
  
}
```

## Single Region Evolution

**GOAL:**

1.  **Define Regions and ROIs:**
    -   Specify the regions of injection (e.g., "DR") and the ROIs to be analyzed (e.g., "Right: PF", "Left: LDT").
2.  **Read Evolution Data:**
    -   For each region, load the data from a CSV file containing evolution metrics.
3.  **Filter and Prepare Data:**
    -   Select columns related to ROIs and mean values.
    -   Filter the data to include only the specified ROIs.
    -   Prepare the data for plotting by transforming from wide to long format, ensuring the order of timepoints is preserved.
4.  **Generate Line Plots:**
    -   For each ROI, create a line plot showing the evolution of synapse density over timepoints.
    -   Customize the plot with labels and a y-axis limit.
5.  **Save Plots:**
    -   Save each plot as a PDF file in a specified directory.
    -   Ensure directory creation if it does not exist and format filenames by cleaning the ROI names.

```{r}
# Regions of injection to plot (even more than one
regions_injections = c("DR")

# ROIs to analyze
ROIs = c("Right: PF", "Left: LDT")

for ( region_injection in regions_injections ){
  
  print(region)
  
  d = data_summary(df, varname="Cell.Density", 
                    groupnames=c("TimePoint", "Region.Injection", "ROI"))
  
  for ( roi in ROIs ){
    
    df_roi = d %>% filter((ROI == roi) & (Region.Injection == region_injection))
    
    # Filter data for the current ROI
    df_roi <- df_roi %>%
      filter(ROI == roi) %>% 
      dplyr::rename(mean = `Cell.Density`)
      
    
    # Line plot using ggplot2
    plot <- ggplot(df_roi, aes(x = TimePoint, y = mean,  group = ROI)) +
            geom_line(color = "blue", size = 1) +
      geom_errorbar(aes(ymin=mean-std, ymax=mean+std), width = 0.2, color = "red") +
      geom_point(color = "blue")+
      coord_cartesian(ylim = c(0, 0.02)) +
      scale_color_brewer(palette="Paired")+
      labs(x = "Timepoint", y = "Mean Synapses Density", title = paste0("Evolution of ROI synapses density\n", roi, "\nInjection Side: ", region))#+#theme_minimal()
    
    print(plot)
    
    #Save the figure
    dir_images = file.path(base_dir, "images", "single_plot_evolution")
    #print(dir_images)
    if (!dir.exists(dir_images)) {
      dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
    }
    # Remove commas and spaces
    roi_clean <- gsub("[, :/]", "", roi)
    
    ggsave(file.path(dir_images, paste0("lineplot_evolution_", roi_clean, "_injection_", region_injection, ".pdf")), plot, width = 5, height = 5)
    
    
  }
}
```

# TODO

## Vulcano Plot

```{r}
# Vulcano plot of 
if (!requireNamespace("DESeq2", quietly = TRUE)) {
  BiocManager::install("DESeq2")
}
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  BiocManager::install("tidyverse")
}

library(DESeq2)
library(tidyverse)
library(ggplot2)

# Choose:
region_injection = "DR"
control_timepoint = "uninjured"
treatment_timepoint = "8 weeks"

#count_data: A matrix of count data (rows are genes, columns are samples).
#col_data: A DataFrame with sample information (including conditions).
#condition: The variable that differentiates between your conditions.

#prepare count_data
count_data = df
count_data = count_data[count_data$`Region.Injection` == region_injection,]
count_data <- count_data %>%
  select(ROI, Synapses, `Brain.ID`) %>%
  pivot_wider(names_from = `Brain.ID`, values_from = Synapses)
count_data = count_data[,-1]

#prepare col_data
col_data = df_meta
col_data = col_data[col_data$`Brain.ID` %in% colnames(count_data),]

# DESeq2 dataset
dds <- DESeqDataSetFromMatrix(countData = count_data,
                              colData = col_data,
                              design = ~ TimePoint)

# Run DESeq analysis
dds <- DESeq(dds)


# Get results
results <- results(dds)

# Print log fold changes and adjusted p-values
#print(results$log2FoldChange)
#print(results$padj)

# Extract results
results_df <- as.data.frame(results)

# Ensure the results contain the necessary columns
#head(results_df)

# Filter out rows with NA values if needed
results_df <- na.omit(results_df)

# Define significance thresholds
significance_threshold <- 0.05
fold_change_threshold <- 1

# Create volcano plot
volcano_plot <- ggplot(results_df, aes(x = log2FoldChange, y = -log10(padj))) +
  geom_point(aes(color = padj < significance_threshold & abs(log2FoldChange) > fold_change_threshold), alpha = 0.5) +
  scale_color_manual(values = c("black", "red")) +
  labs(x = "Log2 Fold Change",
       y = "-Log10 Adjusted P-Value",
       title = "Volcano Plot") +
  #theme_minimal()  +
  #xlim(-2, 2) +
  #ylim(0, 1)+
  geom_vline(xintercept = c(-fold_change_threshold, fold_change_threshold), linetype = "dashed", color = "blue") +
  geom_hline(yintercept = -log10(significance_threshold), linetype = "dashed", color = "blue")+
  theme(legend.position = "none")

# Print the plot
volcano_plot

```

```{r}
# GOAL:
  # Take a region of injection (ex "DR)
  # For each timePoint (ex. "Uninjured")
    # For each ROI 
      # Calculate the mean density across the mouse of the timepoit choosen
  # Take the most "n_roi_displayed" dense region for each time point
    # (so each timepint will have a differt set of ROI)
  # Create a barplot for each time point
    #Put it in the same figure

# Regions of injection to plot (even more than one)
regions_injection = c("DR")

for (region in regions_injection) {  # Iterate over the region_injection
  
  print(region)
  
  # Read df evolution
  df_evolution_region = read_csv(file.path(base_dir, "images", region, "evolution_timeline.csv"))
  
  # Find timepoints of this injection region
  timepoints = df_meta %>% filter(Region.Injection == region) %>% select(TimePoint)
  timepoints = rev(unique(timepoints$TimePoint))
  
  # TimePoint plot
  timepoints_plots = list()
  
  #make a plot for each time point
  for (timepoint in timepoints){
    
    # Name of the mean column and Std in the df
    means_col_name <- paste0("Mean ", timepoint)
    stds_col_name <- paste0("Std ", timepoint)
    
    # Suset to only columns relative to this timeponr
    t = df_evolution_region %>% select(ROI, means_col_name, stds_col_name)
    
    #change name col
    names(t)[names(t) == means_col_name] = "means"
    names(t)[names(t) == stds_col_name] = "stds"

    # Sort the data frame by 'mean' in descending order
    # Select the first most dense regions
    t_sorted <- t %>%
      arrange(desc(means)) %>%
      slice_head(n = n_roi_displayed)
    
    # to force the ROI to be displayed on the x axis on the same order of the most dense
    t_sorted <- t_sorted %>%
      mutate(ROI = factor(ROI, levels = ROI))
    
    #print(t_sorted)
    
    # Create the bar plot with error bars
    plot = ggplot(t_sorted, aes(x = ROI, y = means)) +
      geom_bar(stat = "identity", fill = "lightblue", color = "blue") +
      geom_errorbar(aes(ymin = means - stds, ymax = means + stds), 
                    width = 0.2, color = "red") +
      labs(title = timepoint, x = paste0("First most dense ", n_roi_displayed, " ROI"), y = "Density") +
      coord_cartesian(ylim = c(0, 0.02)) +  # Set y-axis limit
      theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Tilt x-axis labels

    #print(plot)
    
    #put plot in list
    timepoints_plots[[timepoint]] <- plot
  }

  # Print the combined plot
  combined_plot <- wrap_plots(timepoints_plots, ncol = length(timepoints_plots))
  print(combined_plot)
  
  #Save the figure
  dir_images = file.path(base_dir, "images", region)
  #print(dir_images)
  if (!dir.exists(dir_images)) {
    dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
  }
  #ggsave(file.path(dir_images, paste0("barplots_density_different_ROI_", n_roi_displayed, ".pdf")), combined_plot, width = 10, height = 4)


  
```

# Test

```{r}
# Regions of injection to plot (even more than one
regions_injections = c("DR")

# ROIs to analyze
ROIs = c("Right: PF", "Left: LDT")

for ( region_injection in regions_injections ){
  
  print(region)
  
  d = data_summary(df, varname="Cell.Density", 
                    groupnames=c("TimePoint", "Region.Injection", "ROI"))
  
  for ( roi in ROIs ){
    
    df_roi = d %>% filter((ROI == roi) & (Region.Injection == region_injection))
    
    # Filter data for the current ROI
    df_roi <- df_roi %>%
      filter(ROI == roi) %>% 
      dplyr::rename(mean = `Cell.Density`)
      
    
    # Line plot using ggplot2
    plot <- ggplot(df_roi, aes(x = TimePoint, y = mean,  group = ROI)) +
            geom_line(color = "blue", size = 1) +
      geom_errorbar(aes(ymin=mean-std, ymax=mean+std), width = 0.2, color = "red") +
      geom_point(color = "blue")+
      coord_cartesian(ylim = c(0, 0.02)) +
      scale_color_brewer(palette="Paired")+
      labs(x = "Timepoint", y = "Mean Synapses Density", title = paste0("Evolution of ROI synapses density\n", roi, "\nInjection Side: ", region))#+#theme_minimal()
    
    print(plot)
    
    #Save the figure
    dir_images = file.path(base_dir, "images", "single_plot_evolution")
    #print(dir_images)
    if (!dir.exists(dir_images)) {
      dir.create(dir_images, recursive = TRUE) #ensure that all intermediate directories are created as well
    }
    # Remove commas and spaces
    roi_clean <- gsub("[, :/]", "", roi)
    
    ggsave(file.path(dir_images, paste0("lineplot_evolution_", roi_clean, "_injection_", region_injection, ".pdf")), plot, width = 5, height = 5)
    
    
  }
}


```

```{r}
ggplot(df_roi, aes(x = TimePoint, y = mean,  group = ROI)) +
        geom_line(color = "blue", size = 1) +
      geom_errorbar(aes(ymin=mean-std, ymax=mean+std), width = 0.2, color = "red") +
      geom_point(color = "blue")+
      coord_cartesian(ylim = c(0, 0.02)) +
      scale_color_brewer(palette="Paired")+
      labs(x = "Timepoint", y = "Mean Synapses Density", title = paste0("Evolution of ROI synapses density\n", roi, "\nInjection Side: ", region))#+#theme_minimal()
    



```
